Imprecise Datapath Design: An Overclocking Approach

### 2014-05-05 TRETS2013/TRETS_overclocking.tex

\subsection{Accuracy Benefits and Area Overhead in CSA}
We first verify the models for the RCA and CSA in terms of the maximum word-length under the given operating frequencies. For the CSA, the ratio $\mu_{mux}/\mu_c$ can be computed experimentally. We perform post place-and-route simulations on the CSA with 2 stages using Xilinx Virtex-6 FPGA. The delay of the $i^{th}$ stage $d_i$ in~(\ref{Eq:CSA_SingleStageDelay}) is recorded with respect to different word-lengths. The total word-length of CSA can be predicted through~(\ref{Eq:CSA_Timing}). In addition, the maximum word-lengths of the 2-stage CSA and RCA are obtained experimentally by increasing $n_{CSA}$ and $n_{RCA}$ respectively until errors are observed at the output. The comparison between the modeled value and the empirical results is illustrated in Fig.~\ref{Fig_CSA2stage_Timing}. It can be seen that our models for both RCA and CSA match well with the experimental results.
%
%\begin{figure}[htbp]
%        \centering
%        \vspace{-3ex}
%        \subfigure[\scriptsize The modeled value and the experimental results of the maximum word-length of RCA and CSA, with respect to various frequency requirements.]{
%            \includegraphics[width=3.5in]{./Figures/CSA2stage_Model_Timing2.eps}
%            \label{Fig_CSA2stage_Timing}
%        }
%        \subfigure[\scriptsize Hardware resource usage for an RCA and a CSA with 2 and 3 stages.]{
%            \includegraphics[width=3.5in]{./Figures/CSA3stage_Area2.eps}
%            \label{Fig_CSA3stage_Area}
%        }
%        \vspace{-1.5ex}
%    \caption{A comparison between RCA and CSA in terms of the maximum word-length of input signal and the area consumption.}
%\end{figure}
%
\begin{figure}[t]
    %\centering
    \vspace{-3ex}
    \subfigure[]{
        \includegraphics[width=.50\textwidth]{./Figures/CSA2stage_Model_Timing2.eps}
        \label{Fig_CSA2stage_Timing}
    }%
    \subfigure[]{
        \includegraphics[width=.5\textwidth]{./Figures/CSA3stage_Area2.eps}
        \label{Fig_CSA3stage_Area}
    }
    \vspace{-2ex}
    \caption{(a) The modeled value and experimental results of the maximum word-length of RCA and CSA, for various frequency requirements. (b) Hardware resource usage for an RCA and a CSA with 2 and 3 stages.}
    \vspace{-1ex}
\end{figure}

@@

As an example slice through the design space, in Fig.~\ref{Fig_CSA_LUT25} we record the mean value of error at outputs with respect to different operating frequencies for both design scenarios when the number of the available LUTs is set to 25. In this graph, we have labeled the optimum adder structure. Note that in this and all the following experiments within this section, in order to apply our models, the input data is randomly generated following the uniform distribution.
%
\begin{figure}[t]
  \centering
  \vspace{-3ex}
  \includegraphics[width=3.3in]{./Figures/Error_LUT25.eps}
  \vspace{-2ex}
  \caption{A comparison between two design scenarios when the number of available LUT is 25. The RCA and CSA are investigated in the conventional scenario, while the RCA is explored in the proposed new scenario. The results are obtained from post place-and-route simulations on Xilinx Virtex-6 FPGAs.}
  \vspace{-2ex}
  \label{Fig_CSA_LUT25}
\end{figure}

@@

\section{Results and Discussion}\label{Section_Experiments}
\subsection{Case study: FIR filter}
We first assess the accuracy of our proposed models of errors. We record the value of $MRE$ based on (\ref{Eq:MRE}) in a $5^{th}$~order FIR with respect to different operating frequencies. The modeled values of both overclocking error and truncation error of the FIR filter are presented in Fig.~\ref{Fig:FIR} (dotted lines), as well as the actual measurements on the FPGA (solid lines) with two types of input data. The results demonstrate that our models match well with the practical results obtained using the uniform independent inputs.
% FIR
\begin{figure}[tbp]
  \centering
  \vspace{-3ex}
  \includegraphics[width=3.5in]{./Figures/Exp/FIR/FIR_Error_New_TCAD.eps}
  \vspace{-2ex}
  \caption{A demonstration of two design perspectives with a $5^{th}$~order FIR filter, which is implemented on Virtex-6 FPGA. The modeled values of both overclocking errors and truncation errors are presented as dotted lines. The actual FPGA measurements are depicted using solid lines. Two types of inputs are employed in the overclocking scenario: the uniformly distributed data and the real image data from Lena.}
  \label{Fig:FIR}
\end{figure}

>>>>

\subsection{Accuracy Benefits and Area Overhead in CSA}
We first verify the models for the RCA and CSA in terms of the maximum word-length under the given operating frequencies. For the CSA, the ratio $\mu_{mux}/\mu_c$ can be computed experimentally. We perform post place-and-route simulations on the CSA with 2 stages using Xilinx Virtex-6 FPGA. The delay of the $i^{th}$ stage $d_i$ in~(\ref{Eq:CSA_SingleStageDelay}) is recorded with respect to different word-lengths. The total word-length of CSA can be predicted through~(\ref{Eq:CSA_Timing}). In addition, the maximum word-lengths of the 2-stage CSA and RCA are obtained experimentally by increasing $n_{CSA}$ and $n_{RCA}$ respectively until errors are observed at the output. The comparison between the modeled value and the empirical results is illustrated in Fig.~\ref{Fig_CSA2stage_Timing}. It can be seen that our models for both RCA and CSA match well with the experimental results.
%
%\begin{figure}[htbp]
%        \centering
%        \vspace{-3ex}
%        \subfigure[\scriptsize The modeled value and the experimental results of the maximum word-length of RCA and CSA, with respect to various frequency requirements.]{
%            \includegraphics[width=3.5in]{./Figures/CSA2stage_Model_Timing2.eps}
%            \label{Fig_CSA2stage_Timing}
%        }
%        \subfigure[\scriptsize Hardware resource usage for an RCA and a CSA with 2 and 3 stages.]{
%            \includegraphics[width=3.5in]{./Figures/CSA3stage_Area2.eps}
%            \label{Fig_CSA3stage_Area}
%        }
%        \vspace{-1.5ex}
%    \caption{A comparison between RCA and CSA in terms of the maximum word-length of input signal and the area consumption.}
%\end{figure}
%
\begin{figure}[t]
    %\centering
    \vspace{-3ex}
    \subfigure[]{
        \includegraphics[width=.50\textwidth]{./Figures/CSA2stage_Model_Timing_color.eps}
        \label{Fig_CSA2stage_Timing}
    }%
    \subfigure[]{
        \includegraphics[width=.5\textwidth]{./Figures/CSA3stage_Area_color.eps}
        \label{Fig_CSA3stage_Area}
    }
    \vspace{-2ex}
    \caption{(a) The modeled value and experimental results of the maximum word-length of RCA and CSA, for various frequency requirements. (b) Hardware resource usage for an RCA and a CSA with 2 and 3 stages.}
    \vspace{-1ex}
\end{figure}

@@

As an example slice through the design space, in Fig.~\ref{Fig_CSA_LUT25} we record the mean value of error at outputs with respect to different operating frequencies for both design scenarios when the number of the available LUTs is set to 25. In this graph, we have labeled the optimum adder structure. Note that in this and all the following experiments within this section, in order to apply our models, the input data is randomly generated following the uniform distribution.
%
\begin{figure}[t]
  \centering
  \vspace{-3ex}
  \includegraphics[width=3.3in]{./Figures/Error_LUT25_color.eps}
  \vspace{-2ex}
  \caption{A comparison between two design scenarios when the number of available LUT is 25. The RCA and CSA are investigated in the conventional scenario, while the RCA is explored in the proposed new scenario. The results are obtained from post place-and-route simulations on Xilinx Virtex-6 FPGAs.}
  \vspace{-2ex}
  \label{Fig_CSA_LUT25}
\end{figure}

@@

\section{Results and Discussion}\label{Section_Experiments}
\subsection{Case study: FIR filter}
We first assess the accuracy of our proposed models of errors. We record the value of $MRE$ based on (\ref{Eq:MRE}) in a $5^{th}$~order FIR with respect to different operating frequencies. The modeled values of both overclocking error and truncation error of the FIR filter are presented in Fig.~\ref{Fig:FIR} (dotted lines), as well as the actual measurements on the FPGA (solid lines) with two types of input data. The results demonstrate that our models match well with the practical results obtained using the uniform independent inputs.
% FIR
\begin{figure}[tbp]
  \centering
  \vspace{-3ex}
  \includegraphics[width=3.5in]{./Figures/Exp/FIR/FIR_Compare_color.eps}
  \vspace{-2ex}
  \caption{A demonstration of two design perspectives with a $5^{th}$~order FIR filter, which is implemented on Virtex-6 FPGA. The modeled values of both overclocking errors and truncation errors are presented as dotted lines. The actual FPGA measurements are depicted using solid lines. Two types of inputs are employed in the overclocking scenario: the uniformly distributed data and the real image data from Lena.}
  \label{Fig:FIR}
\end{figure}
